%% thesis.tex 2014/04/11
%
% Based on sample files of unknown authorship.
%
% The Current Maintainer of this work is Paul Vojta.

\documentclass{ucbthesis}
\usepackage{amsmath,amssymb,mathtools}  % Better maths support & more symbols
\usepackage{biblatex}
\usepackage{graphicx}% Include figure files
\usepackage{bm} % bold math
\usepackage{braket} % Allows for Dirac notation
\usepackage{nicefrac} % Allows for slanted fractions with sfrac
\usepackage{rotating}

% Multiple references cleanly formatted
\usepackage{cleveref}
\crefname{equation}{equation}{equations}
\crefname{figure}{figure}{figures}

\allowdisplaybreaks[2]% Allow long equations to split pages

% New commands for ChiralPotential

\newcommand{\veff}{V_{12,\text{eff}}}
\newcommand{\rhozero}{\rho_{\text{I}=0}}
\newcommand{\rhoone}{\rho_{\text{I}=1}}

\newcommand{\rhohat}[2]{\hat{\rho}_{\text{I}=#1}\left( #2 \right)}

\newcommand{\yukawa}[1]{\frac{e^{-m_\pi |#1|}}{4\pi |#1|}}
\newcommand{\yukawadimless}[1]{\frac{e^{-m_\pi #1}}{m_\pi #1}}
\newcommand{\yukawanoabs}[1]{\frac{e^{-m_\pi #1}}{4\pi #1}}

\newcommand{\rot}{\vec{r}_{12}}
\newcommand{\rotp}{\vec{r}_{12}\!\!\!'\,\,}

\newcommand{\rotpr}{r_{12}\!\!\!\!'\,\,\,}
\newcommand{\rotphat}{\hat{r}_{12}\!\!\!\!'\,\,\,}

\newcommand{\taudot}{\vec{\tau}_1\cdot\vec{\tau}_2}
\newcommand{\taucrossthree}{\left[\vec{\tau}_1\times\vec{\tau}_2\right]^3}
\newcommand{\tauplusthree}{\frac{\tau_1^3+\tau_2^3}{2}}
\newcommand{\tauminusthree}{\frac{\tau_1^3+\tau_2^3}{2}}

\newcommand{\sigmadot}{\vec{\sigma}_1\cdot\vec{\sigma}_2}
\newcommand{\sigmaplus}{\vec{\sigma}_1+\vec{\sigma}_2}
\newcommand{\sigmaminus}{\vec{\sigma}_1-\vec{\sigma}_2}
\newcommand{\sigmatwo}{[\vec{\sigma}_1\otimes\vec{\sigma}_2]_2}
\newcommand{\sigmaone}{[\vec{\sigma}_1\otimes\vec{\sigma}_2]_1}
\newcommand{\sigmacross}{\vec{\sigma}_1\times\vec{\sigma}_2}

\newcommand{\Gam}[3]{\Gamma_{#1, I=#2}\left(#3\right)}

\newcommand{\fracphantom}{\vphantom{\frac{1}{1}}}

\newcommand{\w}[4]{w_{#1,#2,#3}(#4)}

% New commands for LocalExpansion
\newcommand{\eps}{\epsilon}

% Left acting gradients
\newcommand{\lvec}[1]{\reflectbox{\ensuremath{\vec{\reflectbox{\ensuremath{#1}}}}}}

% \vecrp
%\newcommand{\pvec}[1]{\vec{#1}\mkern2mu\vphantom{#1}}
%\newcommand{\vecrp}{\vec{r}\mkern2mu\vphantom{r}}
\newcommand{\vecrp}{\vec{r}\mkern2mu'}
\newcommand{\hatrp}{\hat{r}\mkern2mu'}

% New commands for SOCoupling

\newcommand{\threej}[6]{ \begin{pmatrix}
  #1 & #2 & #3 \\
  #4 & #5 & #6 
 \end{pmatrix}}

\newcommand{\sixj}[6]{ \begin{Bmatrix}
  #1 & #2 & #3 \\
  #4 & #5 & #6 
 \end{Bmatrix}}

\newcommand{\ninej}[9]{ \begin{Bmatrix}
  #1 & #2 & #3 \\
  #4 & #5 & #6 \\
  #7 & #8 & #9
 \end{Bmatrix}}


% To compile this file, run "latex thesis", then "biber thesis"
% (or "bibtex thesis", if the output from latex asks for that instead),
% and then "latex thesis" (without the quotes in each case).

% Double spacing, if you want it.  Do not use for the final copy.
% \def\dsp{\def\baselinestretch{2.0}\large\normalsize}
% \dsp

% If the Grad. Division insists that the first paragraph of a section
% be indented (like the others), then include this line:
% \usepackage{indentfirst}

%\newtheorem{theorem}{Jibberish}

\bibliography{Introduction/Introduction.bib,ChiralPotential/ChiralPotential.bib,LocalExpansion/LocalExpansion.bib,SOC/SOCoupling.bib}

\hyphenation{mar-gin-al-ia}

\begin{document}

% Declarations for Front Matter

\title{Effective Theories in Few-Body Physics}
\author{Cory Drew Schillaci}
\degreesemester{Fall}
\degreeyear{2015}
\degree{Doctor of Philosophy}
\chair{Professor Wick Haxton}
\othermembers{Professor Joel Moore \\
  Professor Karl van Bibber}
\numberofmembers{3}
\field{Physics}
\campus{Berkeley}

% For a masters thesis, replace the above \documentclass line with
% \documentclass[masters]{ucbthesis}
% This affects the title and approval pages, which by default calls this
% document a "dissertation", not a "thesis".

\maketitle
% Delete (or comment out) the \approvalpage line for the final version.
%\approvalpage
\copyrightpage

\begin{abstract}

Effective theories are a controlled approach to making approximations in physics. We describe here work on two applications of effective theory techniques in few-body physics, specifically to nuclear and atomic systems.

First, we analytically reduce the chiral three-nucleon interaction at N$^2$LO to a density-dependent effective two-body potential by summing the third particle over the states of a spin-symmetric Fermi gas. Results are given for the potential in momentum space and in coordinate space, where the potential is seen to be nonlocal. An expansion of the potential in the difference of the nonlocal coordinates is made in order to arrive at a fully local effective two-body potential. 

We then explore the two-body spectra of spin-$1/2$ fermions in isotropic harmonic traps with external spin-orbit potentials and short range two-body interactions. Using a truncated basis of total angular momentum eigenstates, which is known to be equivalent to an effective theory when the atomic gas is harmonically confined, nonperturbative results are presented for experimentally realistic forms of the spin-orbit coupling: a pure Rashba coupling, Rashba and Dresselhaus couplings in equal parts, and a Weyl-type coupling. The technique is easily adapted to bosonic systems and other forms of spin-orbit coupling.
\end{abstract}


\begin{frontmatter}

%\begin{dedication}
%\null\vfil
%\begin{center}
%To Someone O. Other\\\vspace{12pt}
%Dedication goes here.
%\end{center}
%\vfil\null	
%\end{dedication}

% You can delete the \clearpage lines if you don't want these to start on
% separate pages.

\tableofcontents
\clearpage
\listoffigures
%\clearpage
\listoftables

\begin{acknowledgements}
From the beginning, my parents have always nurtured curiosity about the world around me and held me to high but fair expectations. Many of my earliest memories are of classes at the Pacific Science Center and the Seattle Aquarium where I learned about topics ranging from how viscosity works to the anatomy of squid. When the elementary school librarian thought I wasn't ready for chapter books, a teacher didn't want me to skip a level in math, or anyone didn't believe I was ready for more advanced material than my peers, my parents fought for me and taught me to fight for myself. My mother served as one of my many teachers during my years of homeschooling, provide opportunities to learn and grow in unique ways outside the normal high school system. Of course, their constant example of loving others and providing help to those in need is just as important.

As a graduate student, my advisor Wick Haxton always encouraged me to think about physics intuitively, not just mechanically. Tom Luu, although always too humble, was a great mentor to me through some of the struggles inevitable in the PhD process. Whether meeting in a coffee shop or Skyping half-way around the world, I throughly enjoyed our collaboration. Without him, this thesis would not exist.

Many others have made the journey special outside of the academic aspects. From roasting pigs to enjoying the finer things of life, I have shared great times with my fellow graduate students in addition to the late night homework sessions of our first years. I have been introduced to both the sea and the desert, neither of which I can now live without, and made many new friends in the process. And of course, I met Ana who has shared so many of these adventures with me and supported me through the best and worst of everything.

\end{acknowledgements}

\end{frontmatter}

\pagestyle{headings}

% (Optional) \part{First Part}

\include{Introduction/Introduction}

\include{ChiralPotential/ChiralPotential}

\include{LocalExpansion/LocalExpansion}

\include{SOC/SOCoupling}

\appendix

\include{ChiralPotential/CPAppendix}

\include{LocalExpansion/LEAppendix}

\include{SOC/SOCAppendix}

\printbibliography

\end{document}
